---
title: "rk3110_hw1"
author: "Rose Killian"
date: "1/22/2022"
output: github_document
---

```{r loading libraries and settings}
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Read in .csv

```{r}
hw1_df = 
  read_csv("data/bcdata_Assignment1.csv") %>% 
    janitor::clean_names()
```


# Question 1

Table of quantitative features

```{r}
hw1_df %>% 
  select(age:mcp_1) %>% 
  summary() %>% 
  knitr::kable()
```

# Question 2

Recode BMI

```{r}
q2_df = 
  hw1_df %>% 
  mutate(
    bmi_cat = as_factor(case_when(
      bmi < 16.5 ~ "severely underweight",
      bmi < 18.5 ~ "underweight",
      bmi >= 18.5 & bmi <= 24.9 ~ "normal weight",
      bmi >= 25 & bmi <= 29.9 ~ "overweight",
      bmi >= 30 & bmi <= 34.9 ~ "obesity class I",
      bmi >= 35 & bmi <= 39.9 ~ "obesity class II",
      bmi >= 40 ~ "obesity class III"
    )
  ))

q2_df = 
  q2_df %>% 
  mutate(
    bmi_cat = 
      fct_relevel(bmi_cat,
                  "underweight",
                  "normal weight",
                  "overweight",
                  "obesity class I",
                  "obesity class II"))

#only 5 levels re-ordered as the sample does not contain observations for the lowest or highest BMI categories
```

# Question 3

Bar chart of proportion of breast cancer cases/controls within each BMI category

```{r}
q2_df %>% 
  ggplot(aes(x = bmi_cat)) +
  geom_bar(aes(fill = bmi_cat)) +
  facet_grid(. ~classification) +
  theme(axis.text.x = element_text(angle = 90), legend.position = "none")
```

# Question 4

Logistic regression

```{r}
hw1_df = 
  hw1_df %>% 
  mutate(
    class = as.numeric(classification > 1)
  )

fit_logistic =
  glm(class ~ glucose + homa + leptin + bmi + age, data = hw1_df, family = binomial())

fit_logistic %>% 
  broom::tidy()

confint(fit_logistic, level = 0.95)
```

In order to interpret the beta and 95% confidence interval we must first exponentiate them.

```{r}
exp(cbind(OR = coef(fit_logistic), confint(fit_logistic, level = 0.95)))
```

On average, for every one-unit increase in HOMA-IR, the odds of breast cancer increases by 1.32. We are 95% confident that the true change in odds of breast cancer per unit increase in HOMA-IR is between 0.975 and 1.92.

# Question 5

Linear regression

```{r}
fit = 
  lm(insulin ~ bmi + age + glucose, data = hw1_df)

fit %>% 
  broom::tidy()

confint(fit, level = 0.95)
```

On average, for every one year increase in age, insulin increases by 0.230 units. We are 95% confident that the true increase in insulin for every one year increase in age is between 0.155 and 0.304 units.